---
title: "Assignment 1"
author: "Peter"
date: "2024-02-04"
output: html_document
---

# Car Engine Performance over the years
## This analysis shows the trend of performance in cars over the years


##### (Data set taken from kaggle - https://www.kaggle.com/datasets/rkiattisak/sports-car-prices-dataset )
```{r}
# Importing data from csv into R
Car_data <- read.csv("./Sport car price.csv")
head(Car_data) 
```


```{r}
# Transforming the data set values of column year from character to number
Car_data$Horsepower <- as.numeric(Car_data$Horsepower) 
Car_data$Price..in.USD. <- as.numeric(Car_data$Price..in.USD.)
Car_data$Torque..lb.ft. <- as.numeric(Car_data$Torque..lb.ft.)
```


## Descriptive Statistics
### 1. To find the engine capacity average throughout the years.
### 2. To find the oldest year of the sports car in this dataset, latest year and the mid year
```{r}
# Printing out statistics on the data to get overview insights
summary(Car_data[, 5]) #HP - Quantitative
summary(Car_data[, 3]) #Year - Categorical 
```
# Scatter plot to visualise different types of sports cars manufactured in particular years

```{r}
plot(Car_data$Year, Car_data$Horsepower) #Scatter plot
```

#Plotting the values to find the engine capacity 
```{r}
boxplot(Car_data$Engine.Size..L.) # Box plot

```




